Rewind: Automatically Generated\\Digital Memories from Mobile Geolocation Data

### 2016-09-22 paper/uist_2016/paper.tex
submitted version

\section{Related Work}
% heather - trips vs locations -> (define) semantic memory vs episodic memory - check references, relate to some work done by SenseCam perhaps?
% cut down GPS usage
\subsection{Digital Memories}
Rewind is guided by literature about mediated memories, which provides a theoretical underpinning behind how the Rewind experience is constructed. \textit{Mediated Memories in the Digital Age} offers inspiration informing the relationship between experiences and memories \cite{van2007mediated}. van Dijck notes, ``Remembering is vital to our well being, because without autobiographical memories we would have no sense of past or future, and we would lack any sense of continuity'' (pg. 3). Rewind aims to give the user this sense of continuity in a time segment which has a past and future. By picking a start and end location from a particular trip and generating a Rewind between these points, the user relives their memories in a continuous manner. Similarly in his seminal paper, \textit{As we may think}, Vannevar Bush described \textit{memex}, the envisioned machinery to store infinite information and to allow free, collective access through ``associative trails'', giving mankind the ability of ``...forgetting the manifold things he does not need to have immediately at hand, with some assurance that he can find them again if they prove important''. Rewind in its most ambitious form, perhaps strives to become a potential counterpart of \textit{memex} for retrieving, and sharing of autobiographical memories of one's sensory experiences about past visits.

@@

\subsection{Location History}
Silva and Aizawa build on location data to provide a visual examination of interested traveling routes by linking geo-tagging metadata with online imageries, specifically Google Street View images \cite{deSilva:2009:RMT:1631272.1631414}. The authors proposed a sketch-based interface that would perform spatial and temporal queries on line drawings specified by the users. Rewind resembles this interactive system, in a way that the reconstructed sceneries are of a careful mapping results from geolocation data and Google Street View images. However, Rewind differs in that the mapping is finalized with geolocation data provided by users tracked using smartphones. This means that Rewind serves more of a geteway for recalling episodic memory rather than reviewing a potential travelling itinerary, aiming to provide a mechanism for reminiscence and sharing of past experience while exploring users' behavioral patterns. Moreover, the reconstruction techniques used in Rewind would differentiate the same retrieved image with various temporal cues such as changes in seasons. %; yet the immediate results from the temporal queries suggested by de Silva and Aizawa have not been fully revealed.

Larsen et al. use a spiral chart for a user to see the continuity of their locations, where days are placed around the spiral \cite{larsen2013qs}. This visualization allows users to explore and discover patterns in their daily lives. An early implementation of Visits by Thudt et al. focuses on the map and timeline views of location, interleaving them together \cite{thudt2013visits}. In their visualization, users can move from location to location using their pointer, which at the same time moves back and forth in time. This partially inspired the scrubbing technique in Rewind of traveling back and forth in time by scrubbing left and right, at the same time changing the image of what was seen at that time.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% heather - depth of processing, retrival cues, uniqueness of encoding - Rewind ----> retrival cues
% http://www.psych.utoronto.ca/Neuropsychologylab/PDF/Depth%20of%20processing,%20retrieval%20cues,%20and.pdf
% go into introduction?
% come back to this section if time allows.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Rewind system recreates the visual sensation of travelling down past routes using raw Google Location History data retrieved from location-tracking apps, including the default tracking behavior on Android phones. Differing from previously studied hyperlapse creation techniques, such as the studies by Kopft et al. and Joshi et al \cite{Kopf:2014:FHV:2601097.2601195, Joshi:2015:RHC:2809654.2766954}, Rewind system does not require any cinematic archives from the users, and it is computationally lightweight. Rewind inputs geolocation data, maps out tracked trajectories, and outputs photo-video pairs with Google Street View images (although other street side photo services may be used). The sceneries in each photo-video pair are customized for each user based on times of the day, meteorological conditions and seasonal changes when the route was taken, providing users with multi-dimensional temporal cues. The current version of Rewind system is a open-sourced, web-based application implemented with Google Maps API and third-party JavaScript libraries, and therefore does not transfer sensitive user data nor store them. Rewind can be accessed at [anonymized: url to rewind].  %\ref{fig:flowchart}

@@

Since environmental characteristics, such as meteorological attributes, of images retrieved using Google Street View Static Image API are not directly exposed nor adjustable, it is hard to conclude if the retrieved images would correctly reflect the timely and ambient properties at the time when the user undertook the path. To compensate for this, pixel-wise adjustment was applied on each Google Street View image to better convey time-specific messages including times of the day, seasonal variances and meteorological specifics. Previous work done by \cite{laffont2014transient,shih2013data} have successfully reconstructed images with varying diurnal and seasonal cycles, yet it would be computationally expensive to replicate the proposed techniques in real-time on large sets of images. Considering each Rewind sequence of images would contain two hundred frames, this process would take many hours per route, and it would be challenging to incorporate a large training dataset into the rendering pipeline. Due to the timing constraint, Rewind instead adopts a simplified approach of color space mapping for object segmentation and color adjustment, with the facilitation of CamanJS \cite{caman}.

@@

For meteorological conditions, a NoSQL database populated by data collected from the National Oceanic and Atmospheric Administration (NOAA) weather stations was used for weather lookup (for the study, only data from [anonymized: our state/province] was loaded). Sky colors were adjusted again using the technique mentioned previously for overcast, rainy and snowy sky. For determined rainy and snowy periods of time, an animated precipitation layer was overlaid on the sequence of images and accompanied by corresponding audio effects to emulate the weather conditions exposed by the data (Figure~\ref{fig:weather}). Note that in the user study reported in this paper, an earlier version of the image manipulation was used, which adjusted whole-image color attributes.

>>>>

\section{Related Work}
% heather - trips vs locations -> (define) semantic memory vs episodic memory - check references, relate to some work done by SenseCam perhaps?
% cut down GPS usage
\subsection{Digital Memories}
Rewind is guided by literature about mediated memories, which provides a theoretical underpinning behind how the Rewind experience is constructed. \textit{Mediated Memories in the Digital Age} offers inspiration informing the relationship between experiences and memories \cite{van2007mediated}. van Dijck notes, ``Remembering is vital to our well being, because without autobiographical memories we would have no sense of past or future, and we would lack any sense of continuity'' (pg. 3). Rewind aims to give the user this sense of continuity in a time segment which has a past and future. By picking a start and end location from a particular trip and generating a Rewind between these points, the user relives their memories in a continuous manner. Similarly in his seminal paper, \textit{As we may think}, Vannevar Bush described \textit{memex}, the envisioned machinery to store infinite information and to allow free, collective access, giving mankind the ability of ``...forgetting the manifold things he does not need to have immediately at hand, with some assurance that he can find them again if they prove important''. Rewind in its most ambitious form, perhaps strives to become a potential counterpart of \textit{memex} for retrieving, and sharing of autobiographical memories of one's sensory experiences about past visits.

@@

\subsection{Location History}
Larsen et al. use a spiral chart for a user to see the continuity of their locations, where days are placed around the spiral \cite{larsen2013qs}. This visualization allows users to explore and discover patterns in their daily lives. An early implementation of Visits by Thudt et al. focuses on the map and timeline views of location, interleaving them together \cite{thudt2013visits}. In their visualization, users can move from location to location using their pointer, which at the same time moves back and forth in time. This partially inspired the scrubbing technique in Rewind of traveling back and forth in time by scrubbing left and right, at the same time changing the image of what was seen at that time.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% heather - depth of processing, retrival cues, uniqueness of encoding - Rewind ----> retrival cues
% http://www.psych.utoronto.ca/Neuropsychologylab/PDF/Depth%20of%20processing,%20retrieval%20cues,%20and.pdf
% go into introduction?
% come back to this section if time allows.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Rewind system recreates the visual sensation of travelling down past routes using raw Google Location History data retrieved from location-tracking apps, including the default tracking behavior on Android phones. Differing from previously studied hyperlapse creation techniques, such as the studies by Kopft et al. and Joshi et al \cite{Kopf:2014:FHV:2601097.2601195, Joshi:2015:RHC:2809654.2766954}, Rewind system does not require any cinematic archives from the users, and it is computationally lightweight. Rewind inputs geolocation data, maps out tracked trajectories, and outputs photo-video pairs with Google Street View images (although other street side photo services may be used). The sceneries in each photo-video pair are customized for each user based on times of the day, weather conditions and seasonal changes when the route was taken, providing users with multi-dimensional temporal cues. The current version of Rewind system is a open-sourced, web-based application implemented with Google Maps API and third-party JavaScript libraries, and therefore does not transfer sensitive user data nor store them. Rewind can be accessed at [anonymized: url to rewind].  %\ref{fig:flowchart}

@@

Since environmental characteristics, such as weather attributes, of images retrieved using Google Street View Static Image API are not directly exposed nor adjustable, it is hard to conclude if the retrieved images would correctly reflect the timely and ambient properties at the time when the user undertook the path. To compensate for this, pixel-wise adjustment was applied on each Google Street View image to better convey time-specific messages including times of the day, seasonal variances and weather specifics. Previous work done by \cite{laffont2014transient,shih2013data} have successfully reconstructed images with varying diurnal and seasonal cycles, yet it would be computationally expensive to replicate the proposed techniques in real-time on large sets of images. Considering each Rewind sequence of images would contain two hundred frames, this process would take many hours per route, and it would be challenging to incorporate a large training dataset into the rendering pipeline. Due to the timing constraint, Rewind instead adopts a simplified approach of color space mapping for object segmentation and color adjustment, with the facilitation of CamanJS \cite{caman}.

@@

For weather conditions, a NoSQL database populated by data collected from the National Oceanic and Atmospheric Administration (NOAA) weather stations was used for weather lookup (for the study, only data from [anonymized: our state/province] was loaded). Sky colors were adjusted again using the technique mentioned previously for overcast, rainy and snowy sky. For determined rainy and snowy periods of time, an animated precipitation layer was overlaid on the sequence of images and accompanied by corresponding audio effects to emulate the weather conditions exposed by the data (Figure~\ref{fig:weather}). Note that in the user study reported in this paper, an earlier version of the image manipulation was used, which adjusted whole-image color attributes.
